---
title: "Getting Data into R"
author: "Brendan Apfeld"
date: "October 14, 2016"
documentclass: utbeamerlite
output: beamer_presentation
classoption: aspectratio=1610
fontsize: bigger
---
# A Quick R Refresher
## Packages
* R is super useful, but we supplement its functionality with *packages*
* You install packages (assuming they're in the normal place) using `install.packages("package_name")`
    + You only have to install a package once (though you may need to update it later)
* You have to *load* packages every time you start an R session
    + Load a package in R by using `library(package_name)`
        * Note there are no quotation marks there
    + Loading a package makes all of its functions available to you
    + If you don't need that, you can also do `package_name::function()`
        * E.g. (that we'll see in a second) `rio::import("file_name.csv")`

## Objects and Operators
* Pretty much *everything* in R is an object
    + Which is probably why it's an "OO," or "object-oriented" language
* You "assign" objects to names in R using the ` <- ` operator


# Data Sources
## Online
* Sometimes you'll have a direct link to a csv or txt file
* Other online data may be accessed through an api (and an R package)
* Advantages: potentially easy, "cutting edge data"
* Disadvantages: might change/disappear, often slow

## On Disk
* More often, you'll have some kind of data file on disk
* Advantages: fast, it shouldn't disappear
* Disadvantages: you can break it

# Data Types
## Common Data File Formats
* Spreadsheet-style
    + .xlsx or .xls (Excel)
    + .ods (OpenDocument)
    + Google Sheets
* Proprietary Formats
    + .dta (Stata)
    + .sav (SPSS)
    + .xpt (SAS)
    + .Rdata (R)
* Plain Text Data Formats
    + .csv (our good friend)
    + .dat (not very common)
    + .json
    + .xml

# Getting Data into R
## The (Bad) Old Days
* `read.csv()` or `readr::read_csv()`
* `gdata::read.xls()`
* `foreign::read.dta()` or `haven::read_dta()`
* Which one should I use?
    + Depended on the version of the program used to create the data
    + Each package required different arguments
    + You never really knew what to expect

## One Import Package to Rule them All
Then the `rio` package came along

* Only two commands in the whole package: `rio::import()` and `rio::export()`
* The only argument you *need* for it to work is the file name
* Supports other arguments if you really *want* to change its defaults
* You (almost) never need to change the defaults

## Saving and Opening Rdata Files
* R has its own data format that ends in .Rdata
    + Super compressed
    + Saves and loads quickly
* Save with `save(objects, file = "filename.Rdata")`
* Load with `load("filename.Rdata")`

## Examples with Data on Disk
 Examples in R

## Examples with Data Online
* Can be as simple as 

```wunderground <- read.csv("https://www.wunderground.com/history/airport/ZBAA
/2013/1/1/DailyHistory.html?format=1")```

* The list of possibilities for APIs is far to great to include here
* For an example using the World Bank data API and the corresponding `wbstats` package, check out my [shameless self-promotion](http://brendanapfeld.com/workflow/2016/09/27/wbstats/)

# Visualizing Data
## Visualizing Datasets
* Datasets are often large and unwieldy
* There are a number of ways to easily learn something about them
```{r}
mtcars <- rio::import("mtcars.csv") 
mtcars # a lot of times this is too much for the console
```

## Visualizing Datasets
```{r}
str(mtcars) # look at the structure of the data
```

## Visualizing Datasets
```{r}
names(mtcars) # get the names of the varibles in the dataframe
head(mtcars) # print the first 6 lines
```

## R Also has a Viewer
```{r}
wg <- read.csv("https://www.wunderground.com/history/airport/ZBAA/2013/1/1/DailyHistory.html?format=1")
# View(wg) # commented out because it won't run on a slide
# note the capital "V"
# Can also access by clicking on the object in the 
# environment pane in RStudio
```

# Graphing - Back to R!